---
title: "Untitled"
output: 
  html_document: 
    keep_md: yes
---

This is homework 04 from Lucy Mei.



# Choose my own adventure.
1. Pick one of the data reshaping prompts and do it.
2. Pick a join prompt and do it.

# General data reshaping and relationship to aggregation.

```{r}
library(gapminder)
library(tidyverse)
library(tidyr)
library(dplyr)
```

Prompts: 
## Activity #1
Make my own cheatsheet similar to Tyler Rinker's minimal guide to tidyr.
### First, get my data.
```{r}
country_lifeExp <- gapminder %>%
  select(country, lifeExp, pop) %>%
  group_by(country) %>%
  summarize(mlifeExp = mean(lifeExp), mpop=mean(pop))
  

dat <- data.frame(country_lifeExp)
dat
```

### Gather in Action
```{r}
dat %>% gather (factor, number, -c(country))
```
In this case, we know that number is referred to lifeExp.

### Spread it Back Out
```{r}
dat %>% gather(factor, number, -c(country)) %>%
  spread(factor, number)
```

```{r}
dat %>% gather(factor, number, mlifeExp)
```
In this case, only lifeExp is taken out to reshape.

```{r}
dat %>% gather(factor, number, mlifeExp, mpop)
```

```{r}
dat %>% gather(factor, number, -c(country)) %>%
  spread(country, number)
```
This way, the data can create columns according to each country.

## Activity #2

