---
title: "hm04"
output: 
  html_document: 
    keep_md: yes
---

This is homework 04 from Lucy Mei.



# Choose my own adventure.
1. Pick one of the data reshaping prompts and do it.
2. Pick a join prompt and do it.

# General data reshaping and relationship to aggregation.

```{r}
library(gapminder)
library(tidyverse)
library(tidyr)
library(dplyr)
```

Prompts: 
## Activity #1
Make my own cheatsheet similar to Tyler Rinker's minimal guide to tidyr.
### First, get my data.
```{r}
country_lifeExp <- gapminder %>%
  select(country, lifeExp, pop) %>%
  group_by(country) %>%
  summarize(mlifeExp = mean(lifeExp), mpop=mean(pop))
  

dat <- data.frame(country_lifeExp)
knitr::kable(dat, align = 'c', format = 'markdown', digits = 2)
```

### Gather in Action
In this case, we know that number is referred to lifeExp.
```{r}
dat %>% 
  gather (factor, number, -c(country)) %>%
  knitr::kable(align = 'c', format = 'markdown', digits = 2)
```


### Spread it Back Out
Then I am going to restore the table back to the long format.
```{r}
dat %>% 
  gather(factor, number, -c(country)) %>%
  spread(factor, number) %>%
  knitr::kable(align = 'c', format = 'markdown', digits = 2)
```



I am going to generate a table that only lifeExp is taken out to reshape
```{r}
dat %>% 
  gather(factor, number, mlifeExp) %>%
  knitr::kable(align = 'c', format = 'markdown', digits = 2)
```

```{r}
dat %>% 
  gather(factor, number, mlifeExp, mpop) %>%
  knitr::kable(align = 'c', format = 'markdown', digits = 2)
```


This way, the data can create columns according to each country.
```{r}
dat %>% 
  gather(factor, number, -c(country)) %>%
  spread(country, number) %>%
  knitr::kable(align = 'c', format = 'markdown', digits = 2)
```



## Activity #2 - Make a tibble with one row per year and columns for life expectancy for two or more countries
I want to compare life expectancy between China and India because these two countries are the most populated in Asia. First, I am going to filter out life expectancy from these two countries and make it a new dataframe.
```{r}
coun_life<- gapminder %>%
  filter(country %in% c("China", "India")) %>%
  select(year, country,lifeExp) 
  
dat2 <- data.frame(coun_life)
knitr::kable(dat2, align = 'c', format = 'markdown', digits = 2)
```

This is a table of life expectancy of China and India of each year. I reshaped the data so it is easier to compare the life expectancy between the two countries over time.
```{r}
dat21 <- dat2 %>% 
  spread (country, lifeExp)
knitr::kable(dat21, align = 'c', format = 'markdown', digits = 2, col.names= c("Year", "China", "India"))
```


So, let's make a scatterplot. 
```{r}
ggplot(dat21, aes(year, China)) +
  geom_point(aes(year, China), color = "red") +
  geom_point(aes(year, India), color = "orange") +
  labs(x="Year", y="Life Expectancy", title ="Change of Life Expectancy of China and India Over Time")
```
Life expectancy of China is generally higher compare to that of India except the time around 1960. It is interesting to find out what happened in that year that caused a dramatic drop of life expectancy in China.

## Activity #3 - Compute some measure of life expectancy for all possible combinations of continent and year. Reshape that to have one row per year and one variable for each continent. Or one row per continent and one variable per year.

First, I am going to generate a table with the selected variables I want.
```{r}
dat3 <- gapminder %>%
  select(continent, lifeExp, year) %>%
  group_by(year, continent) %>%
  mutate(meanle = mean(lifeExp))

dat31 <- data.frame(
  dat3 %>%
    select(year, continent, meanle))

knitr::kable(dat31, align = 'c', format = 'markdown', digits = 2, col.names = c("Year", "Continent", "Mean Life Expectancy"))
```

Next, I am going to change it to wide format so I can compare the mean life expectancy between continents for each year.

If I want to plot this table, I will need to plot each continent one by one.
```{r}
dat32 <- dat31 %>%
  reshape2::dcast(year~continent,
                  value.var="meanle",
                  fun.aggregate = mean, na.rm=TRUE)
knitr::kable(dat32, align = 'c', format = 'markdown', digits = 2)
```
 
```{r}
dat32 %>%
  ggplot(aes(year, Africa)) +
  geom_point(aes(year, Africa))+
  geom_point(aes(year, Americas),color="blue") +
  labs(x="Year", y="Mean life expectancy", title = "Change of mean life expectancy of Africa and Americas over time")
```


## Activity #4 - In Window functions, we formed a tibble with 24 rows: 2 per year, giving the country with both the lowest and highest life expectancy (in Asia). Take that table and reshape it so you have one row per year or per year * continent combination.

First, I copied the table.
I am not sure if I need to include the country or not. So I made one that is with the country.
```{r}
dat4 <- gapminder %>%
  select(year, continent, country, lifeExp) %>%
  group_by(continent, year) %>%
  filter(min_rank(desc(lifeExp)) < 2 | min_rank(lifeExp) < 2) %>% 
  arrange(year) 
knitr::kable(dat4, align = 'c', format = 'markdown', digits = 2)
```


```{r}
dat41 <- dat4 %>%
  spread(continent, lifeExp)

knitr::kable(dat41, align = 'c', format = 'markdown', digits = 2)

```

## Activity #5 - Make up a similar set of exercises for myself.
Load the packages.
```{r}
library(reshape2)
library(dplyr)
```

### Group
Create a set of data for the exercise. This table summarize the country and year where the population is lower than the global mean.
```{r}
dat5 <- gapminder %>%
  filter(pop < 2.960e+07) %>%
  select(continent, country, pop, year)

knitr::kable(dat5,align = 'c', format = 'markdown', digits = 2, col.names=c("Continent", "Country", "Population", "Year"))

dat51 <- dat5 %>%
  group_by(continent, country) %>%
  summarize(Occurrence = n(), Sumofpop=sum(pop))
knitr::kable(dat51, align = 'c', format = 'markdown', digits = 2, col.names=c("Continent", "Country", "Number of years with low population", "Sum of population"))
```

### Split
Turn dat51 into dat52.
```{r}
dat52 <- dcast(dat51, continent~country, value.var="Occurrence")
knitr::kable(dat52, align = 'c', format = 'markdown', digits = 2)
```

Now, try to use table.
```{r}
dat53 <-with(dat5, table(continent, country))
knitr::kable(dat53, align = 'c', format = 'markdown', digits = 2)
```

Use dcast to obtain a data.frame.
```{r}
dat54 <-dcast(dat5, continent~country, value.var="pop", fun.aggregate = length)
knitr::kable(dat54, align = 'c', format = 'markdown', digits = 2)
```


### Stack
Turn dat52 into dat55.
```{r}
dat55 <- melt(dat52, id="continent") %>%
  arrange(continent)
knitr::kable(dat55, align = 'c', format = 'markdown', digits = 2, col.names = c("Continent", "Country", "Number of years with low population"))
```


### Join
Create data frames.
```{r}
country <- c("Canada", "Finland", "India")
area <- c(9985000, 338, 3287000)
New_country <- data.frame(country, area)

Old_country <- gapminder %>%
  filter(country %in% c("Canada", "Finland", "India"), year>= 1987, year <=2001) %>%
  select(country, lifeExp)

New_country
Old_country


New_country1 <- data.frame(New_country)
Old_country1 <- data.frame(Old_country)
```
Join by country.
```{r}
join_country <- left_join(Old_country1, New_country1)
knitr::kable(join_country, align = 'c', format = 'markdown', digits = 2)
```

If use merge
```{r}
merge_country <- merge(Old_country1, New_country1)
knitr::kable(merge_country, align = 'c', format = 'markdown', digits = 2)
```

### Subset
To filter out the data I want.
```{r}
subset_new_country <- filter(New_country1, country=="Finland")
knitr::kable(subset_new_country, align = 'c', format = 'markdown', digits = 2)
```

### Transpose
```{r}
New_country1_transpose <- data.frame(t(New_country1))
knitr::kable(New_country1_transpose, align = 'c', format = 'markdown', digits = 2)
```
### Sort
```{r}
country_arrange <- arrange(join_country, country, lifeExp, area)
knitr::kable(country_arrange, align = 'c', format = 'markdown', digits = 2)
```


# Join, merge, look up

## Activity #1 - Create a second data frame, complementary to Gapminder. Join this with Gapminder using a dplyr join function and make some observations about the process and result. Explore the different types of joins.

First, I am going to create a second data frame.
```{r}
country <- c("Canada", "China", "Japan", "United Kingdom", "Germany", "Finland", "Greece","Hungary", "Belgium")
area <- c(9985000, 9597000, 377962, 242495, 357376, 338424,131957, 93030, 30528 )
language <- c("English", "Chinese", "Japanese", "English", "German", "Finnish", "Greek", "Hungarian", "German")

area_language <- data.frame(country, area, language)
knitr::kable(area_language, align = 'c', format = 'markdown', digits = 2)
```

To avoid overwhelming data, I am going to filter some countries from gapminder.
```{r}
gapminder2 <- gapminder %>%
  filter(country %in% c("Canada", "China", "Japan", "United Kingdom", "Germany", "Finland", "Greece","Hungary", "Belgium", "America", "India", "Korea", "France", "Italy"))

knitr::kable(gapminder2, align = 'c', format = 'markdown', digits = 2)
```


Join area_language to gapminder2.
Left_join is joined by country, and two more columns are produced, area and language. For countries not in area_language dataset, NA is displaced in this two columns.
```{r}
left_join(gapminder2, area_language) %>%
  knitr::kable(align = 'c', format = 'markdown', digits = 2)
```

```{r}
semi_join(gapminder2, area_language) %>%
  knitr::kable(align = 'c', format = 'markdown', digits = 2)
```
semi_join only returns things that are in x with a match in y. I think it functions as if I want to filter things in x according to y but the information in y will not be included.

I want to look at the average population density of European countries each year. Here, I used inner_join because inner_join returns things in common between the two data frames.
```{r}
pop_density <- inner_join(gapminder2, area_language) %>%
  filter(continent == "Europe") %>%
  mutate(popden = pop/area)

knitr::kable(pop_density, align = 'c', format = 'markdown', digits = 2)

ggplot(pop_density, aes(year, popden)) +
  geom_point(aes(year, popden, color = country)) +
  geom_line(aes(year, popden, color = country)) +
  theme_bw() +
  labs(x = "Year", y = "Average population density", title = "Change of average population density over time")

```


full_join sums up everything.
```{r}
full_join(gapminder2, area_language) %>%
  knitr::kable(align = 'c', format = 'markdown', digits = 2)
```



anti_join is just the opposite of semi_join.
```{r}
anti_join(gapminder2, area_language) %>%
  knitr::kable(align = 'c', format = 'markdown', digits = 2)
```



## Activity #3 - Explore the base function merge() and match()
```{r}
merge(gapminder2, area_language) %>%
  knitr::kable(align = 'c', format = 'markdown', digits = 2)
```
merge() works very similar to join() function, however, merge() only returns things that are in common. For example, left_join still includes countries that are only in gapminder2 but not area_language.

If we want to make the results of merge() equivalent to that of left_join(), we need add some arguments.
I found the trick from [here](http://zevross.com/blog/2014/04/30/mini-post-for-large-tables-in-r-dplyrs-function-inner_join-is-much-faster-than-merge/).
```{r}
merge(gapminder2, area_language, all.x=TRUE, all.y=FALSE) %>%
  knitr::kable(align = 'c', format = 'markdown', digits = 2)
```



Indeed, it gives the same table.
I think join() has more diveristy that it has more ways of how we want to join the two datasets compare to merge().




Now, let's explore match().
```{r}
a <- knitr::kable(gapminder2, align = 'c', format = 'markdown', digits = 2)
b <- knitr::kable(area_language, align = 'c', format = 'markdown', digits = 2)
match(a,b)
```
match() returns a vector of the positions of the matched item from both datasets. I think it is very different from join() functions since match works with vectors only. 

So one thing I think match() function can do is that, I can extact a column from dataset2, and then put it into dataset1. Here, I extracted *language* column from area_language, then I put it into *pop* in gapminder2, and this is the resulting table I have.
```{r}
gapminder2$pop <- area_language$language[match(gapminder2$country, area_language$country)]

knitr::kable(gapminder2, align = 'c', format = 'markdown', digits = 2)
```











