---
title: "Untitled"
output: 
  html_document: 
    keep_md: yes
---

This is homework 04 from Lucy Mei.



# Choose my own adventure.
1. Pick one of the data reshaping prompts and do it.
2. Pick a join prompt and do it.

# General data reshaping and relationship to aggregation.

```{r}
library(gapminder)
library(tidyverse)
library(tidyr)
library(dplyr)
```

Prompts: 
## Activity #1
Make my own cheatsheet similar to Tyler Rinker's minimal guide to tidyr.
### First, get my data.
```{r}
country_lifeExp <- gapminder %>%
  select(country, lifeExp, pop) %>%
  group_by(country) %>%
  summarize(mlifeExp = mean(lifeExp), mpop=mean(pop))
  

dat <- data.frame(country_lifeExp)
dat
```

### Gather in Action
```{r}
dat %>% 
  gather (factor, number, -c(country))
```
In this case, we know that number is referred to lifeExp.

### Spread it Back Out
```{r}
dat %>% 
  gather(factor, number, -c(country)) %>%
  spread(factor, number)
```

```{r}
dat %>% gather(factor, number, mlifeExp)
```
In this case, only lifeExp is taken out to reshape.

```{r}
dat %>% 
  gather(factor, number, mlifeExp, mpop)
```

```{r}
dat %>% 
  gather(factor, number, -c(country)) %>%
  spread(country, number)
```
This way, the data can create columns according to each country.

## Activity #2 - Make a tibble with one row per year and columns for life expectancy for two or more countries
```{r}
coun_life<- gapminder %>%
  filter(country %in% c("China", "India")) %>%
  select(year, country,lifeExp) 
  
dat2 <- data.frame(coun_life)
dat2
```

This is a table of life expectancy of China and India of each year.
```{r}
dat21 <- dat2 %>% 
  spread (country, lifeExp)
knitr::kable(dat21, align = 'c', format = 'markdown', digits = 2, col.names= c("Year", "China", "India"))
```
It took me a while to figure this out. Year can stay the same, and I will type in what columns I want, then put in the corresponding value.

So, let's make a scatterplot. 
```{r}
ggplot(dat21, aes(year, China)) +
  geom_point(aes(year, China), color = "red") +
  geom_point(aes(year, India), color = "orange") +
  labs(x="Year", y="Life Expectancy", title ="Change of Life Expectancy of China and India Over Time")
```

## Activity #3 - Compute some measure of life expectancy for all possible combinations of continent and year. Reshape that to have one row per year and one variable for each continent. Or one row per continent and one variable per year.

```{r}
dat3 <- gapminder %>%
  select(continent, lifeExp, year) %>%
  group_by(continent, year) %>%
  mutate(meanle = mean(lifeExp))

dat31 <- data.frame(
  dat3 %>%
    select(year, continent, meanle))

dat31
```
```{r}
dat32 <- dat31 %>%
  spread(continent, meanle)
```

## Activity #4 - In Window functions, we formed a tibble with 24 rows: 2 per year, giving the country with both the lowest and highest life expectancy (in Asia). Take that table and reshape it so you have one row per year or per year * continent combination.

First, I copied the table.
```{r}
dat4 <- gapminder %>%
  select(year, continent, country, lifeExp) %>%
  group_by(continent, year) %>%
  filter(min_rank(desc(lifeExp)) < 2 | min_rank(lifeExp) < 2) %>% 
  arrange(year) 
dat4
```
I am not sure if I need to include the country or not. So I made one that is with the country.
```{r}
dat41 <- dat4 %>%
  spread(continent, lifeExp)

knitr::kable(dat41, align = 'c', format = 'markdown', digits = 2)

```

## Activity #5 - Make up a similar set of exercises for myself.
Load the packages.
```{r}
library(reshape2)
library(dplyr)
```

### Group
Create a set of data for the exercise. This table summarize the country and year where the population is lower than the global mean.
```{r}
dat5 <- gapminder %>%
  filter(pop < 2.960e+07) %>%
  select(continent, country, pop, year)

knitr::kable(dat5,align = 'c', format = 'markdown', digits = 2, col.names=c("Continent", "Country", "Population", "Year"))

dat51 <- dat5 %>%
  group_by(continent, country) %>%
  summarize(Occurence = n(), Sumofpop=sum(pop))
knitr::kable(dat51, align = 'c', format = 'markdown', digits = 2, col.names=c("Continent", "Country", "Number of year with low population", "Sum of population"))
```


