---
output: 
  html_document: 
    keep_md: yes
---
#This is STAT545 hw02 from Lucy Mei
# Bring rectangular data in
```{r}
library(gapminder)
library(tidyverse)
```


# Smell test the data
```{r}
typeof(gapminder)
str(gapminder)
```
gapminder is a list?
Its class is data.frame.
There are 6 variables and 1704 rows.
Yes, there are different ways of getting the extend or size. str() can give an overall view of the data. ?
```{r}
str(gapminder)

dim(gapminder)

ncol(gapminder)
nrow(gapminder)
```
The data type of each variable.
```{r}
sapply(gapminder, class)
```
  country continent      year   lifeExp       pop gdpPercap 
 "factor"  "factor" "integer" "numeric" "integer" "numeric" 
 
# Explore individual variables
For country and gdpPercap
The possible values or range of each variable.The variability of gdpPercap is within 0-60000 with most of the countries have gdpPercap between 0-1000. A large proportion of countries are in Africa and similar amount of countries from Americas, Asia and Europe with only fewer than 50 countries from Oceania.
```{r}
hist(gapminder$gdpPercap)
barplot(table(gapminder$continent))
```

# Explore various ploty types
```{r}
ggplot(gapminder, aes(x=lifeExp, y=gdpPercap)) + geom_point(alpha=0.15)
barplot(table(gapminder$continent))
ggplot(filter(gapminder, continent=="Europe"), aes(x=country, y=gdpPercap)) + geom_point(aes(color=country), alpha=0.25)
```

