---
title: "STAT545 hw03"
output: 
  html_document: 
    keep_md: yes
---

## This is hw03 from Lucy Mei.

## Gapminder data
```{r}
library(gapminder)
library(tidyverse)
```

## My mission, high-level
1. Use dplyr and ggplot2 as tools.
2. Make observations about what my tables/figures show and about the process. 
3. Pick at least three of the tasks and make a table and figure.

## Task menu
1. Get the maximum and minimum of GDP per capita for all continents.
```{r}
  gdp <- gapminder %>% group_by(continent) %>% summarize(max_gdpPercap=max(gdpPercap), min_gdpPercap=min(gdpPercap))
  knitr::kable(gdp)
  
```


2. Look at the spread of GDP per capita within the continents.
3. Compute a trimmed mean of life expectancy for different years. Or a weighted mean, weighting by population. Just try something other than the plain vanilla mean.
4. How is life expectancy changing over time on different continents?
5. Report the absolute and/or relative abundance of countries with low life expectancy over time by continent: Compute some measure of worldwide life expectancy - you decide - a mean or median or some other quantile or perhaps your current age. THen determine how many countries on each continent have a life expectancy less than this bench mark, for each year.
6. Find countries with interesting stories. Open-ended and, therefore, hard. Promosing but unsuccessful attempts are encouraged. This will generate interesting questions to follow up on in class.
