---
title: "STAT545 hw03"
output: 
  html_document: 
    keep_md: yes
---

## This is hw03 from Lucy Mei.

## Gapminder data
```{r}
library(gapminder)
library(tidyverse)
```

# My mission, high-level
1. Use dplyr and ggplot2 as tools.
2. Make observations about what my tables/figures show and about the process. 
3. Pick at least three of the tasks and make a table and figure.

# Task menu
## 1. Get the maximum and minimum of GDP per capita for all continents.
```{r results = 'asis'}
  gdp <- gapminder %>% group_by(continent) %>% summarize(max_gdpPercap=max(gdpPercap), min_gdpPercap=min(gdpPercap))
  knitr::kable(gdp)
  ggplot(gdp, aes(continent, max_gdpPercap)) + geom_point(aes(continent, max_gdpPercap,color=continent)) + geom_point(aes(continent, min_gdpPercap, color=continent)) + 
    theme_bw() + labs(x="Continent", y="Max and Min GDP per capita", title="The Maximum and Minimum GDP per capita of all Continents")
```


## 2. Look at the spread of GDP per capita within the continents.

```{r}
p1 <-gapminder %>% select(gdpPercap, lifeExp, continent)
p2 <- ggplot(p1, aes(gdpPercap,lifeExp)) + facet_wrap(~ continent) + geom_point(aes(gdpPercap, lifeExp), color="skyblue3", alpha=0.25) + scale_x_log10()
p2 + theme(strip.background = element_rect(fill="darkolivegreen2"),
           strip.text = element_text(face="bold")) + labs(x="GDP per capita", y="Life Expectancy", title ="The relationship between Life Expectancy and GDP per capita of each continent")
           
```



## 3. Compute a trimmed mean of life expectancy for different years. Or a weighted mean, weighting by population. Just try something other than the plain vanilla mean.
```{r}
gapminder %>%
  filter(continent =="Asia") %>%
  group_by(year) %>%
  summarize(weighted.mean(lifeExp, population))
```



## 4. How is life expectancy changing over time on different continents?
```{r}
l1 <- gapminder %>%
  select(continent, lifeExp, year) %>%
  arrange(year)
l2 <- ggplot(l1, aes(year, lifeExp)) + facet_wrap(~ continent) + geom_point(aes(year, lifeExp), color = "skyblue", alpha=0.25) + geom_smooth(span=0.5, se=FALSE)
l2 + theme_bw() + labs(x="Year", y="Life Expectancy", title = "Change of Life Expectancy over time")
```


5. Report the absolute and/or relative abundance of countries with low life expectancy over time by continent: Compute some measure of worldwide life expectancy - you decide - a mean or median or some other quantile or perhaps your current age. THen determine how many countries on each continent have a life expectancy less than this bench mark, for each year.
6. Find countries with interesting stories. Open-ended and, therefore, hard. Promosing but unsuccessful attempts are encouraged. This will generate interesting questions to follow up on in class.
